---
title: "Project 2 Example 3"
author: "Sadia Perveen"
date: "10/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("stringr", repos = "http://cran.us.r-project.org")
install.packages("tidyr", repos = "http://cran.us.r-project.org")
install.packages("dplyr", repos = "http://cran.us.r-project.org")

require(tidyr)
require(dplyr)
require(stringr)
```
Read in file from source.
```{r}
MALARIA <- readr::read_csv("https://raw.githubusercontent.com/sadia-perveen/Project-2/master/Child-health_malaria_CoD_Feb-2018_WHO_MCEE_-1.csv")
head(MALARIA)

```
Data cleanup:
1. Get rid on first column, its not needed.
2. Update values x2 and x3 row 4 based values in row 5.
3. Drop all columns/rows with na's using drop_na function.
4. Make the first row values in the dataset as the column names.
5. Get rid of 1st two rows using slice.
6. Replace all '-' with 0 using mutate_if function.
```{r}
MALARIA <- select(MALARIA, -c(X1))

MALARIA$X2[4] <- MALARIA$X2[5]
MALARIA$X3[4] <- MALARIA$X3[5]
                            
MALARIA <- MALARIA %>% drop_na()

names(MALARIA) <- as.character(unlist(MALARIA[1,]))

MALARIA <- slice(MALARIA, 3:n())

MALARIA <- mutate_if(MALARIA, 
                is.character, 
                str_replace_all, pattern = "-", replacement = "0")
```

Generate some comparison data, in this case total deaths neanatal and post neonatal based on country for all years. Dividing totals by 10000 to make the numbers readable. 
```{r}
DEATHS_COUNTRY <- MALARIA %>%
  group_by(`Country/area name`) %>%
  summarise(neoNatal = sum(as.integer(gsub(",", "", `Total Neonatal deaths`)))/10000, 
            postNeoNatal = sum(as.integer(gsub(",", "", `Total Post-Neonatal deaths`)))/10000) %>%
  arrange(`Country/area name`)
```

Plot the data in a barplot.
```{r}
barplot(t(as.matrix(DEATHS_COUNTRY[, 2:3])),
        beside = TRUE,
        names.arg = DEATHS_COUNTRY$`Country/area name`,
        legend.text = TRUE,
        las=2, 
        cex.names=.5,
        main = "Deaths NeoNatal/Post Neo Natal By Contry (2000-2016)",
        xlab = "Countries",
        ylab = "# Deaths/10000")
```