---
title: "Project_2_Example_1"
author: "Sadia Perveen"
date: "10/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("stringr", repos = "http://cran.us.r-project.org")
install.packages("tidyr", repos = "http://cran.us.r-project.org")
install.packages("dplyr", repos = "http://cran.us.r-project.org")

require(tidyr)
require(dplyr)
require(stringr)
```
Read in the data from source.
```{r, warning=FALSE}
ELAMAT <- readr::read_csv("https://raw.githubusercontent.com/sadia-perveen/Project-2/master/ELAMAT1.csv")
```

Order data by the mean score in descending order.
```{r, warning=FALSE}
AVG_SCORE <- dplyr::arrange(ELAMAT, desc(as.numeric(MEAN_SCALE_SCORE)))
AVG_SCORE
```

Filter out data that has either a Male or Female listing in SUBGROUP_NAME column and stor eit in ALL_GENDER data frame.
```{r}
ALL_GENDER <- dplyr::filter(ELAMAT, SUBGROUP_NAME == "Female" | SUBGROUP_NAME == "Male")
ALL_GENDER
```

Arrange data in the ALL_GENDER data frame by MEAN_SCALE_SCORE and store it in AVG_SCORE_GEN data frame.
```{r}
AVG_SCORE_GEN <- dplyr::arrange(ALL_GENDER, desc(as.numeric(MEAN_SCALE_SCORE)))
AVG_SCORE_GEN
```

Filter out all Asian or Pacific Islander students and store it in ALL_ASIAN data frame.
Then order the data by MEAN_SCALE_SCORE column.
```{r}
ALL_ASIAN <- dplyr::filter(ELAMAT, SUBGROUP_NAME == "Asian or Pacific Islander")
AVG_SCORE_ASN <- dplyr::arrange(ALL_ASIAN, desc(MEAN_SCALE_SCORE))
AVG_SCORE_ASN
ALL_ASIAN
```

Filter data out based on SUBGROUP_NAME and NRC_DESC.
```{r}
ALL_STUDENTS <- dplyr::filter(ELAMAT, SUBGROUP_NAME == "All Students" &  NRC_DESC == "NYC")
ALL_STUDENTS
ALL_STUDENTS$MEAN_SCALE_SCORE
```

Generate a barplot based on L4_PCT and ITEM_DESC. 
```{r}
barplot(as.numeric(gsub("%", "", ALL_STUDENTS$L4_PCT)), names = ALL_STUDENTS$ITEM_DESC,
        xlab = "Students", ylab = "Percentage",
        main = "PERCENT OF STUDENTS IN LEVEL 4", las=2, cex.names=.5, col=c("darkblue","green"))
```


Generate a barplot based on L1_PCT and ITEM_DESC. 
```{r}
barplot(as.numeric(gsub("%", "", ALL_STUDENTS$L1_PCT)), names = ALL_STUDENTS$ITEM_DESC,
        xlab = "Students", ylab = "Percentage",
        main = "PERCENT OF STUDENTS IN LEVEL 1", las=2, cex.names=.5, col=c("orange","green"))
```